---
title: "Take-home Exercise 1"
description: |
  In this take-home exercise, I will apply the skills learnt in lessons 1 & 2 to create data visualisation by using ggplot2.
author:
  - name: Fei Jiahui 
    url: https://www.linkedin.com/in/jiahui-fei-169841109/
    affiliation: SMU MITB
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Task
There are two data visualization tasks to be completed. They are: 

* A Pareto chart showing the distribution of returns by product sub-category. The superstore-2021.xls file will be used to accomplish this task. 

* A age-sex pyramid showing the demographic structure of Singapore by age cohort and gender. For this task, the data entitle Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2021 will be used. The data set is available at [Department of Statistics](https://www.singstat.gov.sg/) home page.


# Task 1: Pareto Chart

## Installing and loading the required libraries

Before we get started, it is important for us to ensure that the required R packages have been installed. For the purpose of the exercise, the follow tidyverse packages will be used:

* [reaxl](https://readxl.tidyverse.org/) package will be used to read and parse a worksheet into R as a tibble data frame format. (It is important to note that the R object is in tibble data frame and not the generic data frame)

* [dplyr](https://dplyr.tidyverse.org/) package will be used to perform data transformation and data wrangling tasks

* [ggplot2](https://ggplot2.tidyverse.org/) package will be used to building the pareto chart by using the principle of Layered Grammar of Graphs.

The code chunk below is used to check if the necessary R packages are installed in R. If they have yet, then RStudio will install the missing R package(s). If are already been installed, then they will be launch in R environment.


```{r}
packages = c('tidyverse','readxl', 'knitr')
for(p in packages){
  if(!require (p, character.only = T)){
    install.packages(p)
}
library(p,character.only =T)
}
```


## Data Import

In this exercise, superstore-2021 data set will be used. It is an MS Excel workbook. It consists of three worksheets, namely: Orders, People and Returns.

The code chunk below imports superstore-2021.xls into R environment by using [read_csv()](https://readr.tidyverse.org/reference/read_delim.html) function of **readr** package.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
orders <- read_xls("data/Superstore-2021.xls",
                   sheet = "Orders")
returns <- read_xls("data/Superstore-2021.xls",
                   sheet = "Returns")
```

After parsing the worksheet into R, it is a good practice to check the structure and content of the newly tibble data frames in RStudio.


## Data Wrangling

### Joining the two data frames

In this step, the [left_join()](https://dplyr.tidyverse.org/reference/mutate-joins.html) of **dplyr** is used to join the returns data frame and orders data frame by using Order ID as the unique identifier.

```{r}
joined_tab <- left_join(returns, orders,
                        by = c('Order ID' = 'Order ID'))
```

To learn more about Join of dplyr, click on this [link](https://dplyr.tidyverse.org/articles/two-table.html).


### Compute the frequency count by Sub-Category

Next, we are going to compute the frequency count of returns by sub-category. There are two ways to complete the task. The first way is by using the group-by method and the second way is by using the count method of dplyr.

#### The *group-by* method

In the code chunk below, group_by() of dplyr package is used to group the orders by Sub-Category. Then, summarise() of dplyr is used to count (i.e. n()) the number of returned orders.

```{r}
freq_returned <- joined_tab %>%
  group_by(`Sub-Category`) %>%
  summarise('Returns' = n()) %>%
  ungroup()
```

#### The *count* method

The code chunk below shows the alternative way to derive the frequency of returns by sub-category. In this case, count() of dplyr package is used.

```{r}
freq_returned <- joined_tab %>%
  count(`Sub-Category`) %>%
  rename(Returns = n)
```

By default, count() will return a new field called n to store the result. In view of this, rename() of dplyr is used to rename n to Returns.


### Sorting data

Before we can compute the cumulative frequency, we need to sort the values in the sub-category field by the values in the Returns field. To accomplish this task, the arrange() of dplyr package is used as shown in the code chunk below.

```{r}
freq_sorted <- freq_returned %>%
  arrange(desc(Returns))
```


### Computing the cumulative frequency

Out next task is to compute the cumulative frequency of returns by product sub-category. This task will be performed by using mutate() of dplyr package and cumsum() of Base R.

```{r}
freq_cum <- freq_sorted %>%
  mutate(cumfreq = cumsum(Returns))
```


## Plotting the graph

In the code chunk below, geom_col() is used to plot a bar chart by using sub-category as the x-axis and count of returns as the y-axis, while geom_line() is used to draw the line depicting cumulative returns.Note that the scale of the second y axis is calculated using the scale of the first y-axis divided by 552 which is the highest return frequency for a single category.

```{r}
scale <- 552
ggplot(data=freq_cum) +
  geom_col(aes(x = reorder(`Sub-Category`, -Returns), y = `Returns`), 
           color='grey30', 
           fill='lightblue')+
  geom_line(aes(x = reorder(`Sub-Category`, -Returns), y = `cumfreq`*scale/sum(Returns)), 
            color='red', 
            group=1)+
  theme(axis.text.x = element_text(angle = 90))+
  xlab('Sub-Category')+
  scale_y_continuous(name = 'Returns',
                     sec.axis = (~./scale))+
  ggtitle('Distribution of returns by product sub-category')

```


# Task 2: Population Pyramid

## Data Import

In this exercise, *respopagesextod2021* data set will be used. 

The code chunk below imports respopagesextod2021.csv into R environment by using [read_csv()](https://readr.tidyverse.org/reference/read_delim.html) function of **readr** package.

```{r}
pop_data <- read_csv("Data/respopagesextod2021.csv")
```


## Data Wrangling

### Compute population by sex and age group

In the code chunk below, group_by() of dplyr package is used to group the population by gender and age group. Then, summarise() of dplyr is used to count (i.e. n()) the size of population in each group.

```{r}
pop_group <- pop_data %>%
  group_by(`AG`, `Sex`) %>%
  summarise('Population' = sum(`Pop`)) %>%
  ungroup()
```


## Plotting the pyramid

In the code chunk below, geom_col() is used to plot a bar chart by using population as the x-axis and age as the y-axis. An *ifelse* argument is inserted to make the male population inversed so as to creat the pyramid shape.

```{r}
ggplot(data=pop_group, 
       aes(x = ifelse(test = Sex == 'Males', 
                      yes = -Population/1000, 
                      no = Population/1000),
                           y = AG, fill = Sex))+
  geom_col()+
  scale_x_continuous(breaks=seq(-160,160,20),labels = abs(seq(-160,160,20)))+
  labs(x = "Population Size in thousands")
```