---
title: "In-class exercise 5.3"
description: |
  Parallel Coordinates Plot
author:
  - name: Fei Jiahui 
    url: https://example.com/norajones
    affiliation: SMU MITB
    affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Installing and Launching R Packages
The code chunks below are used to install and load the packages in R.

```{r}
packages = c('GGally', 'plotly', 'parcoords', 'parallelPlot', 'tidyverse')

for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```


## Importing World Happiness Data Set
First, important the data into R by using the code below.

```{r}
wh <- read_csv("data/WHData-2018.csv")
```


## Working with ggparcoord() of GGally package
The code chunk below is used to plot a basic parallel coordinates by using ggparcoord() of GGally package.

```{r, fig.height=7, fig.width=7}
ggparcoord(data = wh, 
           columns = c(7:12))
```


Plotting a parallel coordinates with boxplot

```{r, fig.height=7, fig.width=10}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coord. Plot of World Happines Attributes")
```


Parallel coordinates with facet

```{r, fig.height=7, fig.width=10}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax", 
           boxplot = TRUE, 
           title = "Parallel Coord. Plot of World Happines Attributes") +
  facet_wrap(~ Region)
```


## Plotting Interactive Parallel Coordinates: parcoords methods
The code below plot a parallel coordinates plot using the basic parcoords() syntax.

```{r, fig.height=7, fig.width=10}
parcoords(wh[,7:12])
```


Plotting an Interactive Parallel Coordinates Plot

```{r, fig.height=7, fig.width=10}
parcoords(
 wh[,7:12],
 rownames = FALSE,
 alpha = 0.3,
 alphaOnBrushed = 0.2,
 reorderable = T,
 brushMode = '1D-axes',
 width = 700,
 height = 400)
```


## Plotting Interactive Parallel Coordinates Plot: parallelPlot methods
The code chunk below plot an interactive parallel coordinates plot by using parallelPlot().

```{r, fig.height=7, fig.width=10}
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh)
```


Rotate axis label

```{r, fig.height=7, fig.width=10}
parallelPlot(wh,
             rotateTitle = TRUE)
```


Changing the colour scheme

```{r, fig.height=7, fig.width=10}
parallelPlot(wh,
             continuousCS = 'RdBu',
             rotateTitle = TRUE)
```


Parallel coordinates plot with hitogram

```{r, fig.height=7, fig.width=10}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```


